---
title: "Preprocessing of CENTER-TBI dataset"
author:
- Shubhayu Bhattacharyay
- Ari Ercole
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_notebook:
    toc: yes
  html_document:
    toc: yes
    df_print: paged
---
<style type="text/css">
.main-container {
max-width: 1800px;
margin-left: auto;
margin-right: auto;
}
</style>

## I. Initialization

### Import necessary libraries
```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(bestNormalize)
library(gridExtra)
library(latex2exp)
library(caret)
```

### Load data files
```{r message=FALSE}
# List directories of CENTER-TBI datasets
center.TBI.folders <- list.files('../../CENTER-TBI',include.dirs = TRUE)

# Initialize empty lists to store data files
TBI.dataframes <- vector(mode = 'list')
var.lists <- vector(mode = 'list')

# Store all CSVs in the lists
for (i in 1:length(center.TBI.folders)){
  curr.folder <- center.TBI.folders[i]
  var.lists[[curr.folder]] <- read.csv(file.path('../../CENTER-TBI',curr.folder,'variables.csv'))
  TBI.dataframes[[curr.folder]] <- read.csv(file.path('../../CENTER-TBI',curr.folder,'data.csv'))
}

# Isolate IMPACT dataset
impact.dataframe <- TBI.dataframes$sb2406_IMPACT
```

## II. Preprocessing
### Clean CENTER-TBI dataset
```{r message=FALSE}
# Load function to fix IMPACT variable types in the dataframe
source('functions/fix_impact_dataframe.R')

# Drop rows with NA GOSE scores
impact.dataframe <- drop_na(impact.dataframe,GOSE)

# Drop unused columns during imputation
impact.dataframe <- dplyr::select(impact.dataframe,-SiteCode)

# Fix variable types of IMPACT dataframe
impact.dataframe <- fix.impact.dataframe(impact.dataframe)

# Correct erroneous Hemoglobin values (unit discrepancy)
impact.dataframe$Hb[impact.dataframe$Hb >= 50 & !is.na(impact.dataframe$Hb)] <- .1*impact.dataframe$Hb[impact.dataframe$Hb >= 50 & !is.na(impact.dataframe$Hb)]

# Remove patients under the age of 18 from the study
impact.dataframe <- impact.dataframe %>% filter(age >= 18)

# Save cleaned dataset
write.csv(impact.dataframe,'../impact_dataframe.csv',row.names = FALSE)
```

### Observe summary information of CENTER-TBI dataset
```{r, message=FALSE, warning=FALSE}
# Load function to fix IMPACT variable types in the dataframe
source('functions/fix_impact_dataframe.R')

# Load cleaned sample set
impact.dataframe <- read.csv('../impact_dataframe.csv') %>% fix.impact.dataframe()

# Produce table of summary information
knitr::kable(summary(impact.dataframe[,names(impact.dataframe) != 'entity_id']))

# Produce frequency barpolot of GOSE outcome labels
barplot(prop.table(table(impact.dataframe$GOSE)),xlab = 'GOSE',ylab = 'Frequency',main = 'GOSE frequency distribution',ylim = c(0,.5))

# Produce geometric density plots of all numeric IMPACT predictors
impact.dataframe %>%
  keep(is.numeric) %>%                     
  gather() %>%                             
  ggplot(aes(value)) +                     
    facet_wrap(~ key, scales = "free") +   
    geom_density()   
```

## III. Create folds for stratified repeated cross-validation

```{r, message=FALSE, warning=FALSE, eval = FALSE}
# Load function to fix IMPACT variable types in the dataframe
source('functions/fix_impact_dataframe.R')

# Load cleaned sample set labels for stratified cross-validation sampling
impact.dataframe <- read.csv('../impact_dataframe.csv') %>% fix.impact.dataframe()

# Use caret function to create repeated, cross-validation fold partitions:
set.seed(2020)
n.repeats <- 20 # set number of repeats
n.folds <- 5 # set number of folds

testing.folds <- vector(mode = 'list')
for (i in 1:n.repeats){
  curr.repeat.name <- paste0('Repeat',sprintf("%02d",i))
  curr.testing.folds <-
    createFolds(
      as.factor(impact.dataframe$GOSE),
      k = n.folds,
      list = TRUE,
      returnTrain = FALSE
    )
  testing.folds[[curr.repeat.name]] <- curr.testing.folds
}

# Save testing partitions as an RDS file
saveRDS(testing.folds,'../testing_folds.rds')

# Convert testing folds to dataframe format
testing.folds.df <- as.data.frame(matrix(ncol=3,nrow=0))
for (i in 1:length(testing.folds)){
  curr.repeat <- testing.folds[[i]]
  for (j in 1:length(curr.repeat)){
    curr.fold <- curr.repeat[[j]]
    testing.folds.df = rbind(testing.folds.df, data.frame(testIdx = curr.fold, GUPI = impact.dataframe$entity_id[curr.fold], repeat.no = i, fold.no = j))
  }
}
write.csv(testing.folds.df,'../testing_folds.csv', row.names = FALSE)
```

## III. Calculate basic statistical information from clean predictor set
### Table 1: Patient characteristics stratified by GOSE 
```{r, message=FALSE, warning=FALSE, eval = FALSE}
# Load function to fix IMPACT variable types in the dataframe
source('functions/fix_impact_dataframe.R')

# Load IMPACT variables dataframe and fix variable types
impact.dataframe <- read.csv('../impact_dataframe.csv') %>% fix.impact.dataframe()

# Group patients by GOSE scores and measure summary statistics
pt.char.table <-
  impact.dataframe %>% group_by(GOSE) %>% summarise(
    n = n(),
    medAge = median(age),
    Q1Age = quantile(age, 0.25),
    Q3Age = quantile(age, 0.75),
    medGCSm = median(GCSm, na.rm = TRUE),
    Q1GCSm = quantile(GCSm, 0.25, na.rm = TRUE),
    Q3GCSm = quantile(GCSm, 0.75, na.rm = TRUE),
    PercHypoxia = 100 * table(hypoxia)[2] / n(),
    PercHTN = 100 * table(hypotension)[2] / n(),
    medMarshall = median(marshall, na.rm = TRUE),
    Q1Marshall = quantile(marshall, 0.25, na.rm = TRUE),
    Q3Marshall = quantile(marshall, 0.75, na.rm = TRUE),
    PerctSAH = table(tsah)[2] / sum(table(tsah)),
    PercEDH = table(EDH)[2] / sum(table(EDH)),
    meanGlu = mean(glu, na.rm = TRUE),
    sdGlu = sd(glu, na.rm = TRUE),
    meanHb = mean(Hb, na.rm = TRUE),
    sdHb = sd(Hb, na.rm = TRUE),
    notSAH = table(tsah)[2],
    noEDH = table(EDH)[2],
    noHypoxia = table(hypoxia)[2],
    noHypotension = table(hypotension)[2],
    noOneEye = table(unreactive_pupils)[2],
    noTwoEyes = table(unreactive_pupils)[3],
    percOneEye = 100 * table(unreactive_pupils)[2] / sum(table(unreactive_pupils)),
    percTwoEyes = 100 * table(unreactive_pupils)[3] / sum(table(unreactive_pupils))
  )

pt.char.table <- t(pt.char.table)

# Kruskal-Wallis rank sum test for glucose, age, and Hb
glu.kw <- kruskal.test(glu ~ GOSE, data = impact.dataframe)
glu.kw

Hb.kw <- kruskal.test(Hb ~ GOSE, data = impact.dataframe)
Hb.kw

age.kw <- kruskal.test(age ~ GOSE, data = impact.dataframe)
age.kw

# Chi-Squared Tests for number of patients, GCSm, pupillary reactivity, hypoxia, hypotension, Marshall CT classification, tSAH, and EDH
age.chi <- chisq.test(table(impact.dataframe$age,impact.dataframe$GOSE))
gcsm.chi <- chisq.test(table(impact.dataframe$GCSm,impact.dataframe$GOSE))
unreactive_pupils.chi <- chisq.test(table(impact.dataframe$unreactive_pupils,impact.dataframe$GOSE))
hypoxia.chi <- chisq.test(table(impact.dataframe$hypoxia,impact.dataframe$GOSE))
hypotension.chi <- chisq.test(table(impact.dataframe$hypotension,impact.dataframe$GOSE))
marshall.chi <- chisq.test(table(impact.dataframe$marshall,impact.dataframe$GOSE))
tsah.chi <- chisq.test(table(impact.dataframe$tsah,impact.dataframe$GOSE))
EDH.chi <- chisq.test(table(impact.dataframe$EDH,impact.dataframe$GOSE))

# Pairwise Wilcoxon Tests for predictors
age.pw.wx <- pairwise.wilcox.test(impact.dataframe$age, impact.dataframe$GOSE,p.adjust.method = "BH")
glu.pw.wx <- pairwise.wilcox.test(impact.dataframe$glu, impact.dataframe$GOSE,p.adjust.method = "BH")
Hb.pw.wx <- pairwise.wilcox.test(impact.dataframe$Hb, impact.dataframe$GOSE,p.adjust.method = "BH")
GCSm.pw.wx <- pairwise.wilcox.test(impact.dataframe$GCSm, impact.dataframe$GOSE,p.adjust.method = "BH")
marshall.pw.wx <- pairwise.wilcox.test(impact.dataframe$marshall, impact.dataframe$GOSE,p.adjust.method = "BH")
unreactive_pupils.pw.wx <- pairwise.wilcox.test(as.integer(impact.dataframe$unreactive_pupils)-1, impact.dataframe$GOSE,p.adjust.method = "BH")

hypoxia.pw.wx <- pairwise.wilcox.test(as.integer(impact.dataframe$hypoxia)-1, impact.dataframe$GOSE,p.adjust.method = "BH")
hypotension.pw.wx <- pairwise.wilcox.test(as.integer(impact.dataframe$hypotension)-1, impact.dataframe$GOSE,p.adjust.method = "BH")
tsah.pw.wx <- pairwise.wilcox.test(as.integer(impact.dataframe$tsah)-1, impact.dataframe$GOSE,p.adjust.method = "BH")
EDH.pw.wx <- pairwise.wilcox.test(as.integer(impact.dataframe$EDH)-1, impact.dataframe$GOSE,p.adjust.method = "BH")

```

### Spearman correlations with GOSE
```{r, message=FALSE, warning=FALSE, eval = FALSE}
# Load IMPACT dataframe
impact.dataframe <- read.csv('../impact_dataframe.csv')

# Spearman  correlations of predictors with GOSE
age.corr <- cor.test(impact.dataframe$age, impact.dataframe$GOSE, method = 'spearman', exact = FALSE)
GCSm.corr <- cor.test(impact.dataframe$GCSm, impact.dataframe$GOSE, method = 'spearman', exact = FALSE)
unreactive_pupils.corr <- cor.test(impact.dataframe$unreactive_pupils, impact.dataframe$GOSE, method = 'spearman', exact = FALSE)
Hb.corr <- cor.test(impact.dataframe$Hb, impact.dataframe$GOSE, method = 'spearman', exact = FALSE)
glu.corr <- cor.test(impact.dataframe$glu, impact.dataframe$GOSE, method = 'spearman', exact = FALSE)
marshall.corr <- cor.test(impact.dataframe$marshall, impact.dataframe$GOSE, method = 'spearman', exact = FALSE)
```