---
title: "Check assumptions for ordinal and multinomial logistic regression model on CENTER-TBI dataset"
author: 
- Shubhayu Bhattacharyay
- Ari Ercole
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_notebook:
    toc: yes
  html_document:
    toc: yes
    df_print: paged
---
<style type="text/css">
.main-container {
max-width: 1800px;
margin-left: auto;
margin-right: auto;
}
</style>

#### Test assumptions necessary for proportional odds ordinal logistic regression (POLR) and multinomial logistic regression classifiers

## I. Initialization

### Import necessary libraries
```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(MASS)
library(Hmisc)
library(reshape2)
library(caret)
library(Amelia)
library(mice)
library(bestNormalize)
library(UBL)
library(GGally)
library(ggpubr)
library(knitr)
library(brant)
library(car)
library(mlr)
library(rms)
```

### Load imputation information
```{r message=FALSE, warning=FALSE}
# Load MICE object
mi.impact <- readRDS('../mice_impact.rds')

# Identify list of existing imputation directories
imp.dirs <- list.files('../draws',pattern = 'imp*',include.dirs = TRUE,full.names = TRUE)
imp.dirs <- imp.dirs[ file.info(imp.dirs)$isdir ]
```

### Load untransformed, cleaned dataset
```{r message=FALSE, warning=FALSE}
impact.dataframe <- read.csv('../impact_dataframe.csv')
```

## II. Examination of assumptions

### Check for assumption of no multi-collinearity
```{r message=FALSE, warning=FALSE, fig.width=7,fig.height=7}
# Source function to fix IMPACT dataframe
source('./functions/fix_impact_dataframe.R')

# Produce overall correlation plot 
corr.impact.dataframe <- read.csv('../impact_dataframe.csv') %>% 
  fix.impact.dataframe() %>%
  mutate(GCSm = as.factor(GCSm), marshall = as.factor(marshall))
ggpairs(corr.impact.dataframe,columns = c(2,3,5:7,9:13))

# Variance inflation factor (VIF) test
vif.impact.dataframe <- read.csv('../impact_dataframe.csv')
vif.fit <- lm(scale(GOSE) ~ age + unreactive_pupils + GCSm + Hb + glu + hypoxia + hypotension + marshall + tsah + EDH, data = vif.impact.dataframe)
vif.test <- vif(vif.fit) %>% data.frame(variance.inflation.factors=.)
knitr::kable(vif.test)
```

### Check for proportional odds assumption
```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
# Source function to fix IMPACT dataframe
source('./functions/fix_impact_dataframe.R')

# Reload IMPACT dataframe
impact.dataframe <- read.csv('../impact_dataframe.csv') %>% fix.impact.dataframe() %>% mutate(unreactive_pupils = as.integer(unreactive_pupils) - 1)

# Relabel patient type variable
impact.dataframe$PatientType <- factor(impact.dataframe$PatientType,labels = c("ER","Admission","ICU"))

# Examine ratio of log odds of GOSE outcomes across predictor values 
source('functions/log_odds_GOSE.R')
log.odds.table <- with(impact.dataframe, summary(GOSE ~ age + unreactive_pupils + GCSm + Hb + glu + hypoxia + hypotension + marshall + tsah + EDH, fun=log.odds.GOSE, na.include = FALSE))

# Print matrix of log odds ratios
print(log.odds.table)

# Plot log odds ratios across IMPACT predictors
plot(log.odds.table, which = 1:5, xlab = 'logit', main = '', width.factor=1.5,pch=0:14)

# Apply Brant test to statistically test proportional odds assumption
brant.fit <- polr(GOSE ~ age + unreactive_pupils + GCSm + Hb + glu + hypoxia + hypotension + marshall + tsah + EDH, data = impact.dataframe, Hess=TRUE, method = "logistic")

# Print table of Brant test statistics
kable(brant(brant.fit),caption = "Brant Test for Proportional Odds Assumption")
```