---
title: "Examination of Feature Significance in IMPACT prediction models"
author: 
- Shubhayu Bhattacharyay
- Ari Ercole
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
---
<style type="text/css">
.main-container {
max-width: 1800px;
margin-left: auto;
margin-right: auto;
}
</style>

## I. Initialization

### Import necessary libraries
```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(shapper)
library(matrixStats)
library(nnet)
library(MASS)
```

## II. Load trained logistic regression models and extract coefficients and associated p-values

### Load trained logistic regression models and extract coefficients and associated p-values
```{r eval=TRUE, message=FALSE, warning=FALSE}
# Load testing folds for repeated CV
testing.folds <- readRDS('../testing_folds.rds')

# Identify list of existing repeat directories
repeat.dirs <- list.files('../repeated_cv',pattern = 'Repeat*',include.dirs = TRUE,full.names = TRUE)
repeat.dirs <- repeat.dirs[ file.info(repeat.dirs)$isdir ]

# Initialize empty dataframes to store coefficients for the different logistic regression model types
mnlr.coeffs.df <- as.data.frame(matrix(ncol = 7,nrow = 0))
polr.coeffs.df <- as.data.frame(matrix(ncol = 6,nrow = 0))
mnlr.smote.coeffs.df <- as.data.frame(matrix(ncol = 7,nrow = 0))
polr.smote.coeffs.df <- as.data.frame(matrix(ncol = 6,nrow = 0))

# Loop through repeat directories and store relevant information
for (i in 1:length(repeat.dirs)){
  # Identify list of existing fold directories
  fold.dirs <- list.files(repeat.dirs[i],pattern = 'Fold*',include.dirs = TRUE,full.names = TRUE)
  fold.dirs <- fold.dirs[ file.info(fold.dirs)$isdir ]
  # Loop through folds
  for (j in 1:length(fold.dirs)){
    curr.fold.folder <- fold.dirs[j]
    # Load logistic regression model files for current repeat and fold
    curr.mnlr <- readRDS(file.path(curr.fold.folder,'trained_models/mnlr_mdl.rds'))
    curr.mnlr.smote <- readRDS(file.path(curr.fold.folder,'trained_models/mnlr_mdl.rds'))
    curr.polr <- readRDS(file.path(curr.fold.folder,'trained_models/polr_mdl.rds'))
    curr.polr.smote <- readRDS(file.path(curr.fold.folder,'trained_models/polr_mdl.rds'))
    
    # Extract coefficients from current MNLR model
    curr.mnlr.coeff.df <- full_join(as.data.frame(summary(curr.mnlr)$coefficients) %>% 
                                      mutate(GOSE = rownames(.)) %>% 
                                      pivot_longer(cols = !GOSE,names_to = 'predictor',values_to = 'coeff'),
                                    as.data.frame(summary(curr.mnlr)$standard.errors) %>% 
                                      mutate(GOSE = rownames(.)) %>% 
                                      pivot_longer(cols = !GOSE,names_to = 'predictor',values_to = 'coeff.std'),by = c('predictor', 'GOSE')) %>%
      full_join(.,as.data.frame(curr.mnlr$p.values) %>% 
                  mutate(GOSE = rownames(.)) %>% 
                  pivot_longer(cols = !GOSE,names_to = 'predictor',values_to = 'coeff.p.val'),by = c('predictor', 'GOSE')) %>%
      mutate(repeat.idx = i, fold.idx = j)
    
    # Extract coefficients from current MNLR with SMOTE model
    curr.mnlr.smote.coeff.df <- full_join(as.data.frame(summary(curr.mnlr.smote)$coefficients) %>% 
                                      mutate(GOSE = rownames(.)) %>% 
                                      pivot_longer(cols = !GOSE,names_to = 'predictor',values_to = 'coeff'),
                                    as.data.frame(summary(curr.mnlr.smote)$standard.errors) %>% 
                                      mutate(GOSE = rownames(.)) %>% 
                                      pivot_longer(cols = !GOSE,names_to = 'predictor',values_to = 'coeff.std'),by = c('predictor', 'GOSE')) %>%
      full_join(.,as.data.frame(curr.mnlr.smote$p.values) %>% 
                  mutate(GOSE = rownames(.)) %>% 
                  pivot_longer(cols = !GOSE,names_to = 'predictor',values_to = 'coeff.p.val'),by = c('predictor', 'GOSE')) %>%
      mutate(repeat.idx = i, fold.idx = j)
    
    # Extract coefficients from current POLR model
    curr.polr.coeff.df <- data.frame(predictor = names(summary(curr.polr)$coefficients[,1]), 
                                     coeff = summary(curr.polr)$coefficients[,1], 
                                     coeff.std = summary(curr.polr)$coefficients[,2],
                                     coeff.p.val = pnorm(abs(summary(curr.polr)$coefficients[,3]), lower.tail = FALSE) * 2) %>%
      mutate(repeat.idx = i, fold.idx = j)
    
    # Extract coefficients from current POLR with SMOTE model
    curr.polr.smote.coeff.df <- data.frame(predictor = names(summary(curr.polr.smote)$coefficients[,1]), 
                                     coeff = summary(curr.polr.smote)$coefficients[,1], 
                                     coeff.std = summary(curr.polr.smote)$coefficients[,2],
                                     coeff.p.val = pnorm(abs(summary(curr.polr.smote)$coefficients[,3]), lower.tail = FALSE) * 2) %>%
      mutate(repeat.idx = i, fold.idx = j)
    
    # Append current model coefficients to the compiled dataframes
    mnlr.coeffs.df <- rbind(mnlr.coeffs.df,curr.mnlr.coeff.df)
    mnlr.smote.coeffs.df <- rbind(mnlr.smote.coeffs.df,curr.mnlr.smote.coeff.df)
    polr.coeffs.df <- rbind(polr.coeffs.df,curr.polr.coeff.df)
    polr.smote.coeffs.df <- rbind(polr.smote.coeffs.df,curr.polr.smote.coeff.df)
  }
}

# Save compiled coefficient dataframes
write.csv(mnlr.coeffs.df,'../repeated_cv/compiled_coefficients/mnlr_coefficients.csv',row.names = FALSE)
write.csv(mnlr.smote.coeffs.df,'../repeated_cv/compiled_coefficients/mnlr_smote_coefficients.csv',row.names = FALSE)
write.csv(polr.coeffs.df,'../repeated_cv/compiled_coefficients/polr_coefficients.csv',row.names = FALSE)
write.csv(polr.smote.coeffs.df,'../repeated_cv/compiled_coefficients/polr_smote_coefficients.csv',row.names = FALSE)
```

### Load compiled coefficient values from MNLR and POLR and group by clustering variables
```{r eval=TRUE, message=FALSE, warning=FALSE}
options(digits=22)

# Load compiled coefficient dataframes
mnlr.coeffs.df <- read.csv('../repeated_cv/compiled_coefficients/mnlr_coefficients.csv') 
mnlr.coeffs.df$coeff.p.val[mnlr.coeffs.df$coeff.p.val == 0] = .Machine$double.eps
mnlr.coeffs.df <- mutate(mnlr.coeffs.df, z.norm = qnorm(coeff.p.val, lower.tail=FALSE))

mnlr.smote.coeffs.df <- read.csv('../repeated_cv/compiled_coefficients/mnlr_smote_coefficients.csv') 
mnlr.smote.coeffs.df$coeff.p.val[mnlr.smote.coeffs.df$coeff.p.val == 0] = .Machine$double.eps
mnlr.smote.coeffs.df <- mutate(mnlr.smote.coeffs.df, z.norm = qnorm(coeff.p.val, lower.tail=FALSE))

polr.coeffs.df <- read.csv('../repeated_cv/compiled_coefficients/polr_coefficients.csv') %>% 
  mutate(z.norm = qnorm(coeff.p.val, lower.tail=FALSE))

polr.smote.coeffs.df <- read.csv('../repeated_cv/compiled_coefficients/polr_smote_coefficients.csv') %>% 
  mutate(z.norm = qnorm(coeff.p.val, lower.tail=FALSE))

# Group MNLR coefficients by predictor
grouped.mnlr.coeffs.df <- mnlr.coeffs.df %>% group_by(GOSE,predictor) %>% 
  summarise(meanCoeff = mean(abs(coeff)), 
            sdCoeff = sqrt((1/n())*sum(coeff.std^2) +((n()+1)/(n()))*(sd(abs(coeff))^2)), 
            mean.p.value = mean(coeff.p.val),
            std.p.value = sd(coeff.p.val),
            mean.z.norm = mean(z.norm),
            var.z.norm = 1 + (1 + (1/n()))*var(z.norm),
            r.m = (1 + (1/n()))*var(z.norm),
            m = n()) %>%
  mutate(d.o.f. = (m-1)*((1 + (1/r.m))^2)) %>%
  mutate(p.m = pt(q = mean.z.norm, df = d.o.f., lower.tail = FALSE)) %>%
  mutate(formatted = sprintf("%.2f (%.2f)",abs(meanCoeff),sdCoeff), p.value.imp = sprintf("%.2f",p.m))

# Group POLR coefficients by predictor
grouped.polr.coeffs.df <- polr.coeffs.df %>% 
  group_by(predictor) %>% 
  summarise(meanCoeff = mean(abs(coeff)), 
            sdCoeff = sqrt((1/n())*sum(coeff.std^2) +((n()+1)/(n()))*(sd(abs(coeff))^2)), 
            mean.p.value = mean(coeff.p.val),
            std.p.value = sd(coeff.p.val),
            mean.z.norm = mean(z.norm),
            var.z.norm = 1 + (1 + (1/n()))*var(z.norm),
            r.m = (1 + (1/n()))*var(z.norm),
            m = n()) %>%
  mutate(d.o.f. = (m-1)*((1 + (1/r.m))^2)) %>%
  mutate(p.m = pt(q = mean.z.norm, df = d.o.f., lower.tail = FALSE)) %>%
  mutate(formatted = sprintf("%.2f (%.2f)",abs(meanCoeff),sdCoeff), p.value.imp = sprintf("%.2f",p.m))

# Group signficant MNLR coefficients by predictors
pred.group.mnlr.coeffs.df <- grouped.mnlr.coeffs.df %>% group_by(predictor) %>% summarise(absMeanMeanCoeff = (mean(abs(meanCoeff))), sdMeanCoeff = sqrt((1/n())*sum(sdCoeff^2) +((n()+1)/(n()))*(sd(abs(meanCoeff))^2)))%>%
  mutate(formatted = sprintf("%.2f (%.2f)",absMeanMeanCoeff,sdMeanCoeff))

non.significant.mnlr.coeffs.df <- grouped.mnlr.coeffs.df %>% filter(p.m > 0.05)
significant.mnlr.coeffs.df <- grouped.mnlr.coeffs.df %>% filter(p.m <= 0.05)
```

### Load SHAP values from DeepMN and DeepOR and group by clustering variables
```{r eval=TRUE, message=FALSE, warning=FALSE}
# Load compiled SHAP values from DeepMN and DeepOR models
deepMN.shap <- read.csv('../repeated_cv/compiled_shap_values/deepMN_shap_values.csv') %>% 
  pivot_longer(cols = !c(node,repeat.name,fold.name,tune.idx),names_to = "predictor", values_to = "SHAP")
deepOR.shap <- read.csv('../repeated_cv/compiled_shap_values/deepOR_shap_values.csv') %>% 
  pivot_longer(cols = !c(node,repeat.name,fold.name,tune.idx),names_to = "predictor", values_to = "SHAP")

# Group SHAP values by output node and tuning idx
grouped.deepMN.shap <- deepMN.shap %>% 
  group_by(node,tune.idx,predictor) %>% 
  summarise(meanAbsSHAP = mean(abs(SHAP)), sdAbsSHAP = sd(abs(SHAP)),medianAbsSHAP = median(abs(SHAP)), lowerCIAbsSHAP = quantile(abs(SHAP),.25), upperCIAbsSHAP = quantile(abs(SHAP),.75)) %>%
  mutate(formatted = sprintf("%.2f (%.2f)",meanAbsSHAP,sdAbsSHAP)) %>%
  filter(tune.idx == 82) %>%
  pivot_wider(id_cols = predictor, names_from = node, values_from = formatted)

grouped.deepOR.shap <- deepOR.shap %>% 
  group_by(node,tune.idx,predictor) %>% 
  summarise(meanAbsSHAP = mean(abs(SHAP)), sdAbsSHAP = sd(abs(SHAP)),medianAbsSHAP = median(abs(SHAP)), lowerCIAbsSHAP = quantile(abs(SHAP),.25), upperCIAbsSHAP = quantile(abs(SHAP),.75)) %>%
  mutate(formatted = sprintf("%.2f (%.2f)",meanAbsSHAP,sdAbsSHAP)) %>%
  filter(tune.idx == 92) %>%
  pivot_wider(id_cols = predictor, names_from = node, values_from = formatted)

options(digits=5)
greater.wx.test <- pairwise.wilcox.test(deepMN.shap$SHAP, deepMN.shap$predictor,p.adjust.method = "BH",paired = TRUE,alternative = 'greater')
lesser.wx.test <- pairwise.wilcox.test(deepMN.shap$SHAP, deepMN.shap$predictor,p.adjust.method = "BH",paired = TRUE,alternative = 'less')
two.sided.wx.test <- pairwise.wilcox.test(deepMN.shap$SHAP, deepMN.shap$predictor,p.adjust.method = "BH",paired = TRUE)

temp.df.1 <- as.data.frame(greater.wx.test$p.value)
temp.df.1$first.class <- rownames(temp.df.1)
temp.df.1 <- temp.df.1 %>% pivot_longer(cols = -first.class, names_to = 'second.class',values_to = 'p.value') %>% 
  drop_na(p.value) %>% mutate(test = 'greater')

temp.df.2 <- as.data.frame(lesser.wx.test$p.value)
temp.df.2$first.class <- rownames(temp.df.2)
temp.df.2 <- temp.df.2 %>% pivot_longer(cols = -first.class, names_to = 'second.class',values_to = 'p.value') %>% 
  drop_na(p.value) %>% mutate(test = 'less')

temp.df.3 <- as.data.frame(two.sided.wx.test$p.value)
temp.df.3$first.class <- rownames(temp.df.3)
temp.df.3 <- temp.df.3 %>% pivot_longer(cols = -first.class, names_to = 'second.class',values_to = 'p.value') %>% 
  drop_na(p.value) %>% mutate(test = 'two.tailed')

deepOR.greater.wx.test <- pairwise.wilcox.test(deepOR.shap$SHAP, deepOR.shap$predictor,p.adjust.method = "BH",paired = TRUE,alternative = 'greater')
deepOR.lesser.wx.test <- pairwise.wilcox.test(deepOR.shap$SHAP, deepOR.shap$predictor,p.adjust.method = "BH",paired = TRUE,alternative = 'less')
deepOR.two.sided.wx.test <- pairwise.wilcox.test(deepOR.shap$SHAP, deepOR.shap$predictor,p.adjust.method = "BH",paired = TRUE)

deepOR.temp.df.1 <- as.data.frame(deepOR.greater.wx.test$p.value)
deepOR.temp.df.1$first.class <- rownames(deepOR.temp.df.1)
deepOR.temp.df.1 <- deepOR.temp.df.1 %>% pivot_longer(cols = -first.class, names_to = 'second.class',values_to = 'p.value') %>% 
  drop_na(p.value) %>% mutate(test = 'greater')

deepOR.temp.df.2 <- as.data.frame(deepOR.lesser.wx.test$p.value)
deepOR.temp.df.2$first.class <- rownames(deepOR.temp.df.2)
deepOR.temp.df.2 <- deepOR.temp.df.2 %>% pivot_longer(cols = -first.class, names_to = 'second.class',values_to = 'p.value') %>% 
  drop_na(p.value) %>% mutate(test = 'less')

deepOR.temp.df.3 <- as.data.frame(deepOR.two.sided.wx.test$p.value)
deepOR.temp.df.3$first.class <- rownames(deepOR.temp.df.3)
deepOR.temp.df.3 <- deepOR.temp.df.3 %>% pivot_longer(cols = -first.class, names_to = 'second.class',values_to = 'p.value') %>% 
  drop_na(p.value) %>% mutate(test = 'two.tailed')
```