---
title: "Plot figures for the manuscript"
author:
- Shubhayu Bhattacharyay
- Ari Ercole
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_notebook:
    toc: yes
  html_document:
    toc: yes
    df_print: paged
---
<style type="text/css">
.main-container {
max-width: 1800px;
margin-left: auto;
margin-right: auto;
}
</style>

## I. Initialization

### Import necessary libraries
```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(pROC)
library(cvAUC)
library(ggpubr)
library(ROCR)
library(pracma)
library(tidyverse)
library(MASS)
library(Hmisc)
library(reshape2)
library(caret)
library(Amelia)
library(mice)
library(bestNormalize)
library(gridExtra)
library(latex2exp)
library(UBL)
library(GGally)
library(ggpubr)
library(knitr)
library(brant)
library(car)
library(mlr)
library(rms)
library(viridis)
library(shadowtext)
library(VIM)
```

## II. Figure 1: Distributions of numerical and ordinal predictors stratified by GOSE

### (A) Distribution plots (violin and boxplot) of numerical variables
```{r message=FALSE, warning=FALSE, fig.width=6.5, fig.height=2.17}
# Formatted GOSE labels
gose.labels <- c("1", "2 or 3", "4", "5", "6", "7", "8")

# Load untransformed, cleaned dataset
impact.dataframe <- read.csv('../impact_dataframe.csv')

# Pivot IMPACT dataframe into longer dataframe with noncategorical numerical features (including age)
impact.dataframe.long.noncat <- impact.dataframe %>% filter(is.na(glu) | glu <= 20) %>% pivot_longer(cols = c(age,Hb,glu)) %>% drop_na(value) %>% mutate(GOSE = factor(GOSE, order = TRUE)) %>% mutate(GOSE = plyr::mapvalues(GOSE,from = c("1","3","4","5","6","7","8"), to = gose.labels))

# Violin plots of each noncategorical numeric IMPACT predictor against GOSE outcomes
violPlots <- ggplot(impact.dataframe.long.noncat, aes(x = GOSE, y = value, fill = GOSE)) +
  geom_violin(size = .75, 
              alpha = 0.75,
              trim = TRUE) +
  geom_boxplot(width=0.15, 
               fill="white",
               outlier.shape = NA) +
  stat_compare_means(label = "p.signif", 
                     ref.group = ".all.",
                     hide.ns = TRUE) +
  geom_jitter(alpha = 0.15,
              size = .15) +
  facet_wrap(~name, 
             scales = "free_y",
             strip.position = "left",
             labeller = as_labeller(c(age = "Age (y)", glu = "Glucose (mmol/L)", Hb = "Hb (g/dL)") ) ) +
  ylab(NULL) +
  xlab('GOSE at 6 months post-injury') +
  theme_classic() +
  theme(strip.text = element_text(size=20), 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        axis.title.x = element_text(size = 20),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "none",
        aspect.ratio = 1)

print(violPlots)
```

### (B) Segmented bar plots of categorical variables
```{r message=FALSE, warning=FALSE, fig.width=6.5, fig.height=2.17}
# Formatted GOSE labels
gose.labels <- c("1", "2 or 3", "4", "5", "6", "7", "8")

# Load untransformed, cleaned dataset
impact.dataframe <- read.csv('../impact_dataframe.csv')

# Load and fix IMPACT variable dataframe
impact.dataframe.long.cat <- read.csv('../impact_dataframe.csv') %>% pivot_longer(cols = c(GCSm,marshall,unreactive_pupils)) %>% drop_na(value) %>% mutate(GOSE = factor(GOSE, order = TRUE)) %>% mutate(GOSE = plyr::mapvalues(GOSE,from = c("1","3","4","5","6","7","8"), to = gose.labels)) %>% group_by(GOSE, name, value) %>% summarise(count = n())
impact.dataframe.long.cat[impact.dataframe.long.cat$name == 'unreactive_pupils','value']<-impact.dataframe.long.cat[impact.dataframe.long.cat$name == 'unreactive_pupils','value']+1

barPlots <- ggplot(impact.dataframe.long.cat, aes(x = GOSE, y = count, fill = forcats::fct_rev(as.factor(value)))) +
  geom_bar(stat = "identity",
           position = "fill") +
  scale_fill_brewer(palette = "Set2") + 
    facet_wrap(~name, 
             scales = "free_y",
             strip.position = "left",
             labeller = as_labeller(c(GCSm = "Pr(GCSm)",marshall = "Pr(Marshall CT)", unreactive_pupils = "Pr(Unreactive Pupils)") )) +
  ylab('Proportion') +
  xlab('GOSE at 6 months post-injury') +
  theme_classic() +
  theme(strip.text = element_text(size=20), 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "none",
        aspect.ratio = 1)

print(barPlots)
```

## III. Figure 2: True predictor means vs. expected means under the proportional odds assumption

```{r message=FALSE, warning=FALSE, fig.width=6.5, fig.height=7.12}
# Source function to fix IMPACT dataframe
source('./functions/fix_impact_dataframe.R')

# Source function to calculate proportional odds expectations
source('./functions/prop_odds_expectations.R')

# Reload IMPACT dataframe
impact.dataframe <- read.csv('../impact_dataframe.csv') %>% fix.impact.dataframe()

# Formatted GOSE labels
gose.labels <- c("1", "2 or 3", "4", "5", "6", "7", "8")

# `%notin%` <- Negate(`%in%`)
# predictor.set <- names(impact.dataframe)[names(impact.dataframe) %notin% c("entity_id","PatientType","GCS","GOSE")]
# 
# expect.table <- data.frame(matrix(nrow= 0, ncol = 5))
# for (i in 1:length(predictor.set)){
#   curr.pred.name <- predictor.set[i]
#   curr.x <- impact.dataframe[curr.pred.name]
#   not.missing.idx <- which(!is.na(curr.x))
#   non.missing.x <- curr.x[not.missing.idx,1]
#   non.missing.y <- impact.dataframe$GOSE[not.missing.idx]
#   if (is.factor(non.missing.x)){
#     f <- table(non.missing.x)
#     ncat <- length(f)
#     if (ncat < 2) {
#       warning(paste("predictor", curr.pred.name, "only has one level and is ignored"))
#       next
#     }
#     nc <- ncat - 1
#     cats <- (names(f)[order(-f)])[(ncat-nc+1):ncat]
#     for (wcat in cats) {
#       xx <- 1 * (non.missing.x == wcat)
#       curr.expect <- prop.odds.expectation(xx, non.missing.y)
#     }
#   } else {
#     curr.expect <- prop.odds.expectation(non.missing.x, non.missing.y)
#   }
#   curr.expect$predictor <- curr.pred.name
#   expect.table <- rbind(expect.table, curr.expect)
#   names(expect.table) <- names(curr.expect) 
# }
# expect.table <- expect.table %>% 
#   mutate(GOSE = factor(GOSE, order = TRUE)) %>% 
#   mutate(GOSE = plyr::mapvalues(GOSE,from = c("1","3","4","5","6","7","8"), to = gose.labels)) %>%
#   rowwise() %>% 
#   mutate(errorMin = max(xmean.y-xsd.y,0), errorMax = xmean.y+xsd.y)
# 
# write.csv(expect.table,'../prop_odds_expectations.csv',row.names = FALSE)

expect.table <- read.csv('../prop_odds_expectations.csv') %>%
    mutate(GOSE = factor(GOSE, order = TRUE)) 

propOddsPlots <- ggplot(expect.table, aes(x = GOSE, group=1)) +
  geom_line(aes(y = xmean.y), size=1) + 
  geom_point(aes(y = xmean.y), size=3) +
  geom_line(aes(y=xmean.y.po),linetype = "dashed", size=1.20) + 
  facet_wrap(~predictor, 
             ncol = 3,
             scales = "free_y",
             strip.position = "left",
             labeller = as_labeller(c(age = "Age (y)",
                                      unreactive_pupils = "Unreactive Pupils",
                                      hypoxia = "Pr(Hypoxia)",
                                      hypotension = "Pr(Hypotension)",
                                      marshall = "Marshall CT",
                                      tsah = "Pr(tSAH)",
                                      EDH = "Pr(EDH)",
                                      hypotension = "Pr(Hypotension)",
                                      glu = "Glucose (mmol/L)", 
                                      Hb = "Hb (g/dL)",
                                      GCSm = "GCSm")))  +
  ylab(NULL) +
  xlab('GOSE at 6 months post-injury') +
  theme_classic()+
  theme(strip.text = element_text(size=20), 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 16, color = "black"),
        axis.text.y = element_text(size = 16, color = "black"),
        axis.title.x = element_text(size = 20),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size = 2),
        strip.placement = "outside",
        aspect.ratio = 1)

plot(propOddsPlots)
```
## IV. Figure 3: ROC and PRC per each GOSE score at 6 months post-injury

### Load compiled ROC and PRC axes information (calculated during model evaluation)
```{r message=FALSE, warning=FALSE}

# Load IMPACT dataframe and calculate class frequencies for PRC reference
impact.dataframe <- read.csv('../impact_dataframe.csv')
class.freqs <- as.data.frame(table(impact.dataframe$GOSE)/nrow(impact.dataframe)) %>% rename(class = Var1, class.freq = Freq)

# Load and compile axes for each model type
mnlr.plot.roc.pcr.axes <- read.csv('../metrics/mnlr_compiled_plot_roc_prc_axes.csv') %>%
  mutate(Model = 'MNLR')

polr.plot.roc.pcr.axes <- read.csv('../metrics/polr_compiled_plot_roc_prc_axes.csv') %>%
  mutate(Model = 'POLR')

deepMN.plot.roc.pcr.axes <- read.csv('../metrics/deepMN_compiled_plot_roc_prc_axes.csv') %>%
  mutate(Model = 'DeepMN')

deepOR.plot.roc.pcr.axes <- read.csv('../metrics/deepOR_compiled_plot_roc_prc_axes.csv') %>%
  mutate(Model = 'DeepOR')

compiled.plot.roc.pcr.axes <- rbind(mnlr.plot.roc.pcr.axes,polr.plot.roc.pcr.axes,
                                    deepMN.plot.roc.pcr.axes,deepOR.plot.roc.pcr.axes)

# Repair plot endpoints
roc.left.endpoint.idx <- which(compiled.plot.roc.pcr.axes$type == 'roc' & compiled.plot.roc.pcr.axes$x == 0)
roc.right.endpoint.idx <- which(compiled.plot.roc.pcr.axes$type == 'roc' & compiled.plot.roc.pcr.axes$x == 1)
prc.left.endpoint.idx <- which(compiled.plot.roc.pcr.axes$type == 'prc' & compiled.plot.roc.pcr.axes$x == 0)

compiled.plot.roc.pcr.axes[roc.left.endpoint.idx,c("mean.y","lower.ci.y","upper.ci.y")] <- 0
compiled.plot.roc.pcr.axes[roc.right.endpoint.idx,c("mean.y","lower.ci.y","upper.ci.y")] <- 1
compiled.plot.roc.pcr.axes[prc.left.endpoint.idx,c("mean.y","lower.ci.y","upper.ci.y")] <- 1

# Change class integers to GOSE labels
compiled.plot.roc.pcr.axes <- compiled.plot.roc.pcr.axes %>%
  mutate(class = as.factor(class)) %>%
  left_join(class.freqs,by = 'class') %>%
  mutate(class = plyr::mapvalues(class, 
                                 from = c(1,3,4,5,6,7,8),
                                 to = c("GOSE: 1", "GOSE: 2 or 3", "GOSE: 4", "GOSE: 5", "GOSE: 6", "GOSE: 7", "GOSE: 8")))

# Change order of models
compiled.plot.roc.pcr.axes$Model <- factor(compiled.plot.roc.pcr.axes$Model, levels = c('MNLR','POLR','DeepMN','DeepOR'))
```

### (A) Receiver operating characteristic (ROC) curves per each GOSE score at 6-mo post-injury
```{r message=FALSE, warning=FALSE, fig.width=6.5, fig.height=3.4}
roc.curves <- compiled.plot.roc.pcr.axes %>% 
  filter(type == 'roc') %>%
  ggplot(aes(x = x)) +
  facet_wrap( ~ class,
              scales = 'free',
              ncol = 4) +
  xlab("False Positive Rate") +
  ylab("True Positive Rate") +
  coord_cartesian(ylim = c(0,1),xlim = c(0,1))+
  geom_ribbon(aes(ymin = lower.ci.y, ymax = upper.ci.y, fill = Model), alpha = 0.3) +
  geom_line(aes(y = mean.y, color = Model), alpha = 0.75, size=1.20) +
  guides(linetype = FALSE, color = guide_legend(nrow = 2)) +
  geom_segment(x = 0, y = 0, xend = 1, yend = 1,alpha = 0.5,linetype = "dashed",size=1, color = 'gray70')+
  theme_classic()+
  theme(
    strip.text = element_text(size=20), 
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 12, color = "black"),
    axis.text.y = element_text(size = 12, color = "black"),
    axis.title.x = element_text(size = 22),
    axis.title.y = element_text(size = 22),
    strip.background = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size = 2),
    legend.position = c(0.85, .25),
    legend.title = element_text(size=20),
    legend.text=element_text(size=16),
    plot.title = element_text(hjust = 0.5),
    aspect.ratio = 1
  )

plot(roc.curves)
```

### (B) Precision-recall (PRC) curves per each GOSE score at 6-mo post-injury
```{r message=FALSE, warning=FALSE, fig.width=6.5, fig.height=3.4}
prc.curves <- compiled.plot.roc.pcr.axes %>% 
  filter(type == 'prc') %>%
  ggplot(aes(x = x)) +
  facet_wrap( ~ class,
              scales = 'free',
              ncol = 4) +
  xlab("Recall") +
  ylab("Precision") +
  coord_cartesian(ylim = c(0,1),xlim = c(0,1))+
  geom_ribbon(aes(ymin = lower.ci.y, ymax = upper.ci.y, fill = Model), alpha = 0.3) +
  geom_line(aes(y = mean.y, color = Model), alpha = 0.75, size=1.20) +
  guides(linetype = FALSE, color = guide_legend(nrow = 2)) +
  geom_segment(aes(y = class.freq, yend = class.freq),x = 0, xend = 1, alpha = 0.5,linetype = "dashed",size=1, color = 'gray70')+
  theme_classic()+
  theme(
    strip.text = element_text(size=20), 
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 12, color = "black"),
    axis.text.y = element_text(size = 12, color = "black"),
    axis.title.x = element_text(size = 22),
    axis.title.y = element_text(size = 22),
    strip.background = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size = 2),
    legend.position = c(0.85, .25),
    legend.title = element_text(size=20),
    legend.text=element_text(size=16),
    plot.title = element_text(hjust = 0.5),
    aspect.ratio = 1
  )

plot(prc.curves)
```

## V. Figure 4: Probability calibration curves per each GOSE score at 6 months post-injury

```{r message=FALSE, warning=FALSE, fig.width=6.5, fig.height=3.4}
# Set GOSE panel labels
gose.labels <- c("GOSE: 1", "GOSE: 2 or 3", "GOSE: 4", "GOSE: 5", "GOSE: 6", "GOSE: 7", "GOSE: 8")
# 
# # Load compiled calibration curves
# compiled.cal.curves.df <- read.csv('../metrics/compiled_calib_curves.csv') %>%
#   mutate(class = factor(class)) %>%
#   mutate(class = plyr::mapvalues(class, from = c(1,3,4,5,6,7,8), to = gose.labels))
# 
# # Interpolate calibration curves into common x-axes values
# interp.cal.curves.df <- data.frame(matrix(ncol = ncol(compiled.cal.curves.df),nrow=0))
# xq <- seq(from = 0, to = 1, length.out = 21)
# for (curr.label in unique(compiled.cal.curves.df$class)){
#   for (curr.bs.idx in unique(compiled.cal.curves.df$bs_idx)){
#     for (curr.model in unique(compiled.cal.curves.df$Model)){
#       curr.idx <- which(compiled.cal.curves.df$class == curr.label & compiled.cal.curves.df$bs_idx == curr.bs.idx & compiled.cal.curves.df$Model == curr.model)
#       if (length(curr.idx) == 0){
#         next
#       } else if (length(curr.idx) == 1){
#         interp.cal.curves.df <- rbind(interp.cal.curves.df, data.frame(prob_pred = xq[which.min(abs(compiled.cal.curves.df$prob_pred[curr.idx] - xq))], prob_true = compiled.cal.curves.df$prob_true[curr.idx], Model = curr.model, bs.idx = curr.bs.idx, class = curr.label))
#         next
#       }
#       interpol.object <- approx(x = compiled.cal.curves.df$prob_pred[curr.idx],y = compiled.cal.curves.df$prob_true[curr.idx],xout = xq)
#       interp.cal.curves.df <- rbind(interp.cal.curves.df, data.frame(prob_pred = interpol.object$x, prob_true = interpol.object$y, Model = curr.model, bs.idx = curr.bs.idx, class = curr.label))
#     }
#   }
# }
# # Save interpolated calibration dataframe
# write.csv(interp.cal.curves.df,'../metrics/interp_compiled_calib_curves.csv')
# 
# plot.interp.cal.curves.df <- interp.cal.curves.df %>%
#   group_by(class, Model, prob_pred) %>%
#   summarise(mean.y = mean(prob_true,na.rm = TRUE),
#             lower.ci.y = quantile(prob_true,.025,na.rm = TRUE),
#             upper.ci.y = quantile(prob_true,.975,na.rm = TRUE))
# 
# # Save summarized, interpolated calibration dataframe
# write.csv(plot.interp.cal.curves.df,'../metrics/plot_interp_compiled_calib_curves.csv')

# Load summary dataframes of interpolated ROC axes
plot.interp.cal.curves.df <- read.csv('../metrics/plot_interp_compiled_calib_curves.csv')

# Change order of models
plot.interp.cal.curves.df$Model <- factor(plot.interp.cal.curves.df$Model, levels = c('MNLR','POLR','DeepMN','DeepOR'))

calib.plot <-
  plot.interp.cal.curves.df %>%
  ggplot(aes(x = prob_pred)) +
  facet_wrap( ~ class,
              scales = 'free',
              ncol = 4) +
  xlab("Mean Predicted Probability") +
  ylab("Fraction of Positives") +
  geom_ribbon(aes(ymin = lower.ci.y, ymax = upper.ci.y, fill = Model),
              alpha = 0.3) +
  guides(linetype = FALSE, color = guide_legend(nrow = 2)) +
  geom_segment(x = 0, y = 0, xend = 1, yend = 1,alpha = 0.5,linetype = "dashed",size=1, color = 'gray70')+
  geom_line(aes(y = mean.y, color = Model), alpha = 0.8, size=1.20) +
  coord_cartesian(ylim = c(0,1),xlim = c(0,1))+
  theme_classic()+
  theme(
    strip.text = element_text(size=22), 
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 12, color = "black"),
    axis.text.y = element_text(size = 12, color = "black"),
    axis.title.x = element_text(size = 22),
    axis.title.y = element_text(size = 22),
    strip.background = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size = 2),
    legend.position = c(0.85, .25),
    legend.title = element_text(size=20),
    legend.text=element_text(size=16),
    plot.title = element_text(hjust = 0.5),
    aspect.ratio = 1
  )

plot(calib.plot)
```

## VI. Figure 5: Normalised confusion matrices of each model type

```{r message=FALSE, warning=FALSE, fig.width=6.5, fig.height=7.66}
# Load deepMN confusion matrix axes and average across repeats
deepMN.cm.values <- read.csv('../metrics/deepMN_confusionMatrices.csv') %>%
  group_by(true_labels,predicted_labels) %>%
  summarise(mean_cm_prob = mean(cm_prob), sd_prob = sd(cm_prob)) %>%
  mutate(Model = 'DeepMN')

# Load deepOR confusion matrix axes and average across repeats
deepOR.cm.values <- read.csv('../metrics/deepOR_confusionMatrices.csv') %>%
  group_by(true_labels,predicted_labels) %>%
  summarise(mean_cm_prob = mean(cm_prob), sd_prob = sd(cm_prob)) %>%
  mutate(Model = 'DeepOR')

# Load mnlr confusion matrix axes and average across repeats
mnlr.cm.values <- read.csv('../metrics/mnlr_confusionMatrices.csv') %>%
  group_by(true_labels,predicted_labels) %>%
  summarise(mean_cm_prob = mean(cm_prob), sd_prob = sd(cm_prob)) %>%
  mutate(Model = 'MNLR')

# Load polr confusion matrix axes and average across repeats
polr.cm.values <- read.csv('../metrics/polr_confusionMatrices.csv') %>%
  group_by(true_labels,predicted_labels) %>%
  summarise(mean_cm_prob = mean(cm_prob), sd_prob = sd(cm_prob)) %>%
  mutate(Model = 'POLR')

# Compile all model types into one dataframe
compiled.confusion.matrix <- rbind(deepMN.cm.values, deepOR.cm.values, mnlr.cm.values, polr.cm.values)
compiled.confusion.matrix <- compiled.confusion.matrix %>% mutate(Model = factor(Model, levels = c("MNLR","POLR","DeepMN","DeepOR")),
                                                                  true_labels = factor(true_labels),
                                                                  predicted_labels = factor(predicted_labels))

# Plot confusion matrices
confusionMatrixPlots <- ggplot(compiled.confusion.matrix, aes(x = predicted_labels,y = true_labels,fill = mean_cm_prob))+
  geom_tile() +
  geom_shadowtext(aes(label= sprintf("%0.2f",mean_cm_prob)),color="white", size = 6)+
  scale_fill_viridis(discrete=FALSE,limits=c(0, 1), breaks=seq(0,1,by=0.25)) +
  guides(fill = guide_colourbar(barwidth = 40,
                                title = 'Mean proportion of predicted label given true label',
                                title.position = 'top',
                                title.hjust = .5)) +
  ylab(label = "True Labels") +
  xlab(label = 'Predicted Labels')+
  scale_y_discrete(limits = rev(levels(compiled.confusion.matrix$true_labels)))+
  facet_wrap(~Model, nrow = 2, ncol = 2)+
  theme_classic()+
  theme(
    strip.text = element_text(size=22, color = 'black'), 
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 18, color = "black",angle = 45, hjust = 1, vjust = 1),
    axis.text.y = element_text(size = 18, color = "black",angle = 45, hjust = 1, vjust = 0),
    axis.title.x = element_text(size = 22),
    axis.title.y = element_text(size = 22),
    strip.background = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size = 2),
    legend.position = 'bottom',
    legend.title = element_text(size = 18, color = "black"),
    legend.text=element_text(size=16),
    aspect.ratio = 1
  )

plot(confusionMatrixPlots)
```

## VII. Supplementary Figure 1: Missingness pattern of IMPACT predictor variables in CENTER-TBI

```{r message=FALSE, warning=FALSE, fig.width=5}
# Load function to fix IMPACT variable types in the dataframe
source('functions/fix_impact_dataframe.R')

# Load cleaned sample set labels for stratified cross-validation sampling
impact.dataframe <- read.csv('../impact_dataframe.csv') %>% fix.impact.dataframe()

# Convert pupillary reactivity variable type to integer for imputation
impact.dataframe$unreactive_pupils <- as.integer(impact.dataframe$unreactive_pupils) - 1

# Remove non-IMPACT predictors
plot.impact.dataframe <- impact.dataframe %>% dplyr::select(-c(entity_id,PatientType,GCS,GOSE))

# Shorten plot labels to fit
plt.labels <- names(plot.impact.dataframe)
plt.labels[1] <- "Age"
plt.labels[2] <- "P.R."
plt.labels[5] <- "Glu."
plt.labels[6] <- "Hypoxia"
plt.labels[7] <- "HoTN"
plt.labels[8] <- "Marshall"
plt.labels[9] <- "tSAH"

# Produce both barplots of missing variables and combinations plot
miss.aggr <- aggr(plot.impact.dataframe,numbers=TRUE,labels = plt.labels)
```
## VIII. Supplementary Figure 2: Effect of optimized Box-Cox transformation and standardization on distributions of numerical and ordinal variables
```{r message=FALSE, warning=FALSE, fig.width=6.5, fig.height=4.5}
# Load function to fix IMPACT variable types in the dataframe
source('functions/fix_impact_dataframe.R')

# Load IMPACT dataframe and fix variable types
impact.dataframe <- read.csv('../impact_dataframe.csv') %>% fix.impact.dataframe()

# Identify columns to undergo Box-Cox normalization
bc.columns <- c('age','GCSm','Hb','glu','marshall')

# # Initialize empty lists to store Box-Cox models and make modifiable copy of dataset
# bc <- vector(mode = 'list')
# bc.impact.dataframe <- impact.dataframe
# copy.impact.dataframe <- impact.dataframe
# 
# # Initialize empty list to store ggplot objects
# bc.plots <- list()
# # Loop through to-normalize columns
# for (curr.col in bc.columns){
#   if (curr.col %in% c('GCSm','marshall')){
#     copy.impact.dataframe[[curr.col]] <- as.factor(copy.impact.dataframe[[curr.col]])
#     if (curr.col == 'marshall'){
#       copy.impact.dataframe[[curr.col]] <- plyr::mapvalues(copy.impact.dataframe[[curr.col]], from = c('1','2','3','4','5','6'), to = c('I','II','III','IV','V','VI'))
#     }
#     bc.plots[[paste0(curr.col, '.og.density')]] <- copy.impact.dataframe %>%
#       drop_na(curr.col) %>%
#       ggplot(aes_string(x = curr.col)) +
#       geom_bar(aes(y = ..prop.., group = 1)) +
#       theme_classic() +
#       theme(axis.text.x = element_text(size = 16, color = "black"),
#             axis.text.y = element_text(size = 16, color = "black"),
#             axis.title.x = element_blank(),
#             axis.title.y = element_blank(),
#             panel.border = element_rect(colour = "black", fill=NA, size = 2),
#             legend.position = "none")
#   } else {
#     bc.plots[[paste0(curr.col, '.og.density')]] <- copy.impact.dataframe %>%
#       drop_na(curr.col) %>%
#       ggplot(aes_string(x = curr.col)) +
#       geom_histogram(aes(y =..density..), bins = min(100,length(unique(copy.impact.dataframe[[curr.col]]))) ) +
#       theme_classic() +
#       theme(axis.text.x = element_text(size = 16, color = "black"),
#             axis.text.y = element_text(size = 16, color = "black"),
#             axis.title.x = element_blank(),
#             axis.title.y = element_blank(),
#             panel.border = element_rect(colour = "black", fill=NA, size = 2),
#             legend.position = "none")
#   }
#   # First, plot original density of current numeric variable
# 
#   # Transform current numeric variable and store transformed variable
#   bc[[curr.col]] <- boxcox(impact.dataframe[[curr.col]],
#                            standardize = TRUE)
#   bc.impact.dataframe[, curr.col] <- bc[[curr.col]]$x.t
#   curr.shapiro.test <- shapiro.test(bc.impact.dataframe[[curr.col]])
# 
#   # Second, plot transformed density of current numeric variable
#   bc.plots[[paste0(curr.col, '.tf.density')]] <- ggplot(bc.impact.dataframe, aes_string(x = curr.col)) +
#     geom_density() +
#     theme_classic() +
#     theme(axis.text.x = element_text(size = 16, color = "black"),
#           axis.text.y = element_text(size = 16, color = "black"),
#           axis.title.x = element_blank(),
#           axis.title.y = element_blank(),
#           panel.border = element_rect(colour = "black", fill=NA, size = 2),
#           legend.position = "none")
#   # Third, plot normal q-q plot of transformed numeric variable
#   bc.plots[[paste0(curr.col, '.tf.normqq')]] <- ggplot(bc.impact.dataframe, aes_string(sample = curr.col)) +
#     stat_qq() +
#     stat_qq_line() +
#     theme_classic() +
#     theme(axis.text.x = element_text(size = 16, color = "black"),
#           axis.text.y = element_text(size = 16, color = "black"),
#           axis.title.x = element_blank(),
#           axis.title.y = element_blank(),
#           panel.border = element_rect(colour = "black", fill=NA, size = 2),
#           legend.position = "none")
#   # Fourth, plot the transformation (Box-Cox) function of the current numeric variable
#   xx <- seq(min(impact.dataframe[[curr.col]],na.rm = TRUE), max(impact.dataframe[[curr.col]],na.rm = TRUE), length = 500)
#   yy <- predict(bc[[curr.col]],newdata = xx)
#   xrng <- range(xx)
#   yrng <- range(yy)
#   curr.tf.df <- data.frame(xx,yy)
#   bc.plots[[paste0(curr.col,'.tf.function')]] <- ggplot(curr.tf.df,aes(xx,yy)) +
#     geom_line() +
#     theme_classic() +
#     theme(axis.text.x = element_text(size = 16, color = "black"),
#           axis.text.y = element_text(size = 16, color = "black"),
#           axis.title.x = element_blank(),
#           axis.title.y = element_blank(),
#           panel.border = element_rect(colour = "black", fill=NA, size = 2),
#           legend.position = "none") 
#     # annotate("text", x=xrng[2], y=yrng[1], label=TeX(sprintf("$\\lambda = %0.2f$", bc[[curr.col]]$lambda)),hjust = 1, vjust = 0, size = 8)
# }
# # Save Box-Cox plots object:
# saveRDS(bc.plots,'../box_cox_plots.rds')

# Load Box-Cox plots object:
bc.plots <-readRDS('../box_cox_plots.rds')
do.call("grid.arrange",c(bc.plots,nrow=length(bc.columns),ncol=4))

print('Y-axis labels (Top-to-bottom): Age, GCSm, Hb, Glucose, Marshall CT')
print('X-axis labels (Left-to-Right): Original Density, Transformed Density, Transformed Density, Transformed Q-Q, Transformation Function')
```
